---
title: "DATA 608 - Module 1"
author: "Sarah Wigodsky"
date: "August 27, 2018"
output: html_document
---

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```
```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

It is helpful to find the standard deviation of the quantitative variables.  

Standard deviation in rank:
```{r}

sd(inc$Rank)

```

Standard deviation in growth rate:
```{r}

sd(inc$Growth_Rate)

```
Standard deviation in revenue:
```{r}

sd(inc$Revenue)

```

Number of data points:
```{r}
nrow(inc)
```


## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r distribution-of-companies, eval=TRUE, fig.width=4.5, fig.height=11}
x <- barplot(sort(summary(inc$State)), cex.axis=1.0, cex.names=.80, horiz=TRUE, col="lightblue", las=2, main="Distribution of Companies by State", ylab="State", xlab="Number of Companies")
text(x, labels=sort(summary(inc$State)), pos=4)
```

## Question 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

New York is the state with the 3rd most companies. When initially creating a plot, some of the outliers were so far afield (particularly for business products & services and consumer products & services) that it was challenging to see trends.  I therefore changed the scale so that we can only view the number of employees up until 1100. The following graph shows the median number of employees by industry sorted from greatest to smallest.  

```{r employees-in-industries-in-NY, eval=TRUE, fig.height=12,fig.width=13, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
ny_industry <- inc[complete.cases(inc),] %>%
  filter(State=='NY') %>%
  group_by(Industry) 

ggplot(ny_industry, aes(x=reorder(Industry, Employees, FUN=median), y=Employees)) +
  geom_boxplot(aes(color=reorder(Industry, Employees, FUN=median))) + 
  coord_flip(ylim=c(0,1100)) +
  labs(x = "Industry", title = "Median Number of Employees by Industry in New York", color="Industry")
```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.  

The following graph shows the median revenue per employee by industry, organized by median from greatest to smallest.  When initially creating a plot, some of the outliers were so far afield (particularly for energy, manufacturing and software) that it was challenging to see trends.  I therefore changed the scale so that we can only view the revenue per employees up until $2,000,000.  


```{r revenue-per-employee, eval=TRUE, fig.height=12,fig.width=13, warning=FALSE}
industry <- inc[complete.cases(inc),] %>%
  mutate(rev_per_emp = Revenue/Employees) %>%
  group_by(Industry) 

ggplot(industry, aes(x=reorder(Industry, rev_per_emp, FUN = median), y=rev_per_emp)) +
  geom_boxplot(aes(color=reorder(Industry, rev_per_emp, FUN = median))) + 
  coord_flip(ylim = c(0,2000000)) +
  labs(x = "Industry", y = "Revenue per Employee", title = "Median Revenue Per Employee", color="Industry")
```